---
title: "Project_III"
author: "Patricia Kiambo"
date: "6/30/2019"
output: html_document
---
## Malaria Cases by Counties in Kenya - Year: 2006 & 2013

## Data Import
  Data was sourced from https://data.humdata.org/dataset/bed-nets-and-illness-by-county-kenya

## Formulate the question
 I am interested in comparing the malaria cases in 2006 and 2013

## Read in the data
```{r}
library(dplyr)

malaria_ke <- read.csv("malaria_explore.csv")
```

Is the difference of means significant. Choose a categorical variable to create the two samples and take the means of a quantitative variable 

```{r}
## Boxplot to 
boxplot(malaria_ke$Malaria_2006,  malaria_ke$Malaria_2013,col="sky blue")

## One Sample t-test
t.test(malaria_ke$Malaria_2006,malaria_ke$Malaria_2013)

## Two Sample t-test
t.test(x=malaria_ke$Malaria_2006,y=malaria_ke$Malaria_2013, 
       mu=.16, 
       alt="two.sided", 
       conf=.95, 
       var.eq=F, 
       paired=F)

```
 
Justify why you reject or fail to reject the Null hypothesis based on p-value and confidence interval 

```{r}
## p-value = 0.9749
## confidence level = 95%
## A small p-value (typically ≤ 0.05) indicates strong evidence against the null hypothesis, so I reject the null hypothesis.

```

