---
title: "Project_II"
author: "Patricia Kiambo"
date: "6/23/2019"
output: html_document
---
## Malaria Cases by Counties in Kenya - Year: 2006 & 2013

## Data Import
  Data was sourced from https://data.humdata.org/dataset/bed-nets-and-illness-by-county-kenya
  
## EDA Checklist

## Formulate your question
 I am interested in comparing the malaria cases in 2006 and 2013, and also exploring the correlation of poverty rate to malaria cases in 2006.

## Read in your data
```{r}
library(dplyr)
library(readr) 
malaria_ke <- read.csv("malaria_explore.csv")
```

## Check the packaging
```{r}
nrow(malaria_ke)
ncol(malaria_ke)
```

## Run str()
```{r}
str(malaria_ke)
class(malaria_ke)
```

## Look at the top and bottom of the data
```{r}
head(malaria_ke[, c(2:5,7)])

tail(malaria_ke[, c(2:5,8)])
```

## Check the n's (the count)
```{r}
 table(malaria_ke$Health_Facilities)

## How many countries are represented in this dataset?
select(malaria_ke, County) %>% unique %>% nrow 

## Names of Counties in Kenya
unique(malaria_ke$County) 
```

## Validate with at least one external data source
```{r}
## 
summary(malaria_ke$Malaria_2006)

## The Department of Health, Kenya attempts to keep the cases of Malaria under .45%. In 2006, the hightest malaria case was 0.6330. The mean barely meets the recommendation.

## Let us look at the quantiles (deciles) of the data: 
quantile(malaria_ke$Malaria_2006, seq(0, 1, 0.1)) 

mtest<-filter(malaria_ke, Malaria_2006 > 0.35)
nrow(mtest)
head(mtest)
```

## Find an easy solution
## 
```{r}
## Which counties have the highest cases of Malaria in 2006
ranking <- group_by(malaria_ke, County) %>%  summarize(malaria_ke = mean(Malaria_2006)) %>% as.data.frame %>%  arrange(desc(malaria_ke))
head(ranking, 10)
```

## Challenge the solution
```{r}
## Which counties had a reduced malaria cases by 2013?
ranking <- group_by(malaria_ke, County) %>%  summarize(malaria_ke = (Malaria_2013-Malaria_2006)) %>% as.data.frame %>%  arrange(malaria_ke)
tail(ranking, 10)
```

## Follow up
```{r}
## Calculates the 

```

## Exploratory Graphics

```{r}
library(ggplot2)

## Histogram
ggplot(malaria_ke, aes(Malaria_2006)) +
  geom_histogram(binwidth = .025)

ggplot(malaria_ke, aes(Malaria_2006)) + 
  geom_histogram(aes(y=..density..), color="red", fill="green", binwidth = .025) +
  geom_density(fill="gold",  alpha=.4)

## Scatterplots
with(malaria_ke,plot(Poverty_Rate_2006, Health_Facilities,xlab="Poverty Rate 2006", ylab = "Health_Facilities"))
     
ggplot(data = malaria_ke) + 
  geom_point(mapping = aes(x = Poverty_Rate_2006, y = Health_Facilities, 
                           color = Under_BedNet_2006,size= HealthSpending_2006))

```
